### 2017-02-24 Paper/Paper_Eric.tex
Added discussion of baseline summary, improved calibration discussion and introduction of analysis part - mentioning the importance of analyzing welfare implications for both groups (u,ue)

In our presentation we have discussed welfare implications of changes in the unemployment policy. We have done so using a simple Aiyagari framework, where agents are subject to exogenous unemployment shocks. These shocks are assumed to be independent of the policy itself. Furthermore, the unemployment rate, which only depends on the different transition probabilities is so too. In other words, changes of the level unemployment insurance do not affect the unemployment rate. This strongly contradicts empirical literature.\footnote{See \citep{decker} for a review.} The main argument being that higher unemployment benefits will decrease incentives for the workers to look for jobs. Furthermore, there is a great deal of theoretical literature, consisting of search-match and moral hazard models suggesting the same. (cite some guys)

@@

I will start by discussing the model then go on by describing how I conducted the welfare analysis. In the thrid part I present my results after a brief summary of the results presented during our presentation - the baseline case with exogenous transition probabilities. Finally, I conclude with a critical discussion.

@@

For our presentation we have taken two different forms of measurement into account - Consumption equivalent and Cash equivalent. I will here only discuss the Cash Equivalent. However, the conclusions I will draw from my analysis are the same when using the Consumption Equivalent (See Appendix 1).\footnote{That being said, the two measures differ in the manner they weight different groups. The Consumption equivalent does weight different level's of consumption differently. A concave utility function implies that a consumption change of the same order is weighted differently by rich individuals with a high consumption to poor individuals with a low consumption.}

@@

In the following section I will analyse the importance of considering disincentive and incentive effects when considering changes in the unemployment benefits. Within the model described above and built de foundation for our results presented in class, the transition probabilities and the unemployment rate are exogenously determined. We therefore abstracted from such incentive effects. In the following experiment I will loosen this strong assumption and analyze the implications for the welfare analysis.

After I introduce the Calibration, I proceed with a short summary of the results obtained in the baseline case and then proceed with a more detailed discussion of the case with an endogenous job-finding probability.

@@

I will use the KrusellandSmith calibration, which founded the basis of our analysis for the presentation. However, since I am not considering aggregate risk, I will use the job-finding probability of 0.4\footnote{For the presentation we took $\pi_{ue}=0.35$. Note that the results for the welfare analysis in the baseline case do not change.See appendix 1}, which corresponds to an average unemployment duration of 2.5 quarters or 32.5 weeks. Note that this a rather high estimate of unemployment duration.\footnote{\cite{mukoyama} makes a calibration for the post 1980 US economy, has a job-finding probability of 0.26 corresponding to 17 weeks of unemployment, thus providing a better match for the real data.}
Other variables are, a discount factor $\beta = 0.99$, productivity $z = 1$, $\alpha = 0.36$, depreciation rate $\theta = 0.025$, the baseline employment $rate l=0.9$, the baseline replacement rate $\mu = 0.15$.

@@

When we conducted our prior welfare analysis we found that different groups in society want different policies - namely, the unemployed would prefer to change to higher unemployment benefits and the employed would prefer a change to a less distributive system. Fruthermore, we also found that generally agents do not want full unemployment insurance, although this would clearly eliminate idiosyncratic risk. 
Figure 1 illustrates this well:

@@

>>>>

In our presentation we have discussed welfare implications of changes in the unemployment policy. We have done so using a simple Aiyagari framework, where agents are subject to exogenous unemployment shocks. These shocks are assumed to be independent of the policy itself. Furthermore, the unemployment rate, which only depends on the different transition probabilities is so too. In other words, changes of the level of unemployment insurance do not affect the unemployment rate. This strongly contradicts empirical literature.\footnote{See \citep{decker} for a review.} The main argument being that higher unemployment benefits will decrease incentives for the workers to look for jobs. Furthermore, there is a great deal of theoretical literature, consisting of search-match and moral hazard models suggesting the same. (cite some guys)

@@

Sections 1 and 2 discuss the model and the welfare analysis, respectively. In section 3 a brief summary of the results presented in class is followed by the experiment, which takes changes in the job-finding probability into account. Finally, section 5 concludes.

@@

For our presentation we have considered two different forms of welfare measurements - the Consumption equivalent and the Cash equivalent. I will here only discuss the Cash Equivalent. However, the conclusions I will draw from my analysis are the same when using the Consumption Equivalent (See Appendix 1).\footnote{That being said, the two measures differ in the manner they weight different groups. The Consumption equivalent does weight different level's of consumption differently. A concave utility function implies that a consumption change of the same order is weighted differently by rich individuals with a high consumption to poor individuals with a low consumption.}

@@

In the following section I will analyse the importance of considering job-search effects when considering changes in the unemployment benefits. How does the welfare analysis change when taking into account that the unemployment duration rises in the benefit level? \\
In order to provide a more detailed welfare analysis, I will look at both the employed and the unemployed when I compare the two cases. Although society on average may clearly favour one particular outcome, it may still be reasonable to implement a certain policy to help the more vulnerable to the detriment of the majority. Typically, the unemployed benefit relatively more from higher benefits than the employed. The latter being in the larger majority, thus leads to the result being biased towards them. 
\\ \\
After I introduce the Calibration, I proceed with a short summary of the results obtained in the baseline case, which corresponds to the results discussed in class.
I then conduct an experiment on the basis of empirical results, which show that an increase in the replacement rate leads to an increase in the unemployment duration. Using this relationship I will recalculate the welfare effects, letting the job-finding probability fluctuate with the benefit level. The second welfare analysis is then compared to the baseline case.

@@

The calibration used here follows \cite{DenHaan20101} is the same as in the presentation, with the small difference, that the transition matrix corresponds to the case with no aggregate risk. The job-finding probability $\pi_{ue}$ therefore takes is slightly higher and takes on the value 0.4 instead of 0.35, which would correspond to the case of a recession when including aggregate risk.\footnote{Since we were also considering aggregate risk in our presentation, we had to chose between the recession and boom state for the Aiyagari calibration and chose the latter. Note, however, that this has no impact on the qualitative result of the welfare analysis.} The calibration is based on quarterly data matching microeconomic data or long-run model considerations. A job-finding probability of 0.4, thus, corresponds to an average unemployment duration of 2.5 quarters or 32.5 weeks. \footnote{Note that here we overestimate the average unemployment duration, since agents remain unemployed for at least one quarter.E.g. \cite{mukoyama} uses a $\pi_{ue}$ leading to an average duration of 17 weeks, which corresponds to post US 1980 economy data.} The discount factor is set at $\beta = 0.99$, productivity at $z = 1$, the capital share at $\alpha = 0.36$, and the depreciation rate $\delta = 0.025$. Finally, the baseline unemployment rate corresponds to  $1- l=0.1$ and the baseline replacement rate is $\mu = 0.15$.

@@

When we conducted our prior welfare analysis we found that different groups in society want different policies - namely, the unemployed would prefer to change to higher unemployment benefits and the employed would prefer a change to a less redistributive system. Fruthermore, we also found that generally agents do not want full unemployment insurance, although this would clearly eliminate idiosyncratic risk. 
Figure 1 illustrates this well:

@@

As benefits increase taxes rise to meet the higher demand. Moreover, higher insurance decreases idiosyncratic risk and thus, precautionary savings decrease. This in turn reduces aggregate capital leading to an increase in the return rate on capital and a decrease in wages. 
The employed, thus, experience a sharp decrease in their net-wage, but profit more from higher return rates on capital. The first clearly dominates.\footnote{ (mukoymama working paper.shows that welfare effects are mainly determined by the redistributional effects of higher taxes and higher benefits. He, however, considers lump-sum taxes and therefore cannot show a different effect for unemployed and employed.} Consequentially the employed prefer a reduction in benefits compared to the baseline level. 
Contrary to the employed, the unemployed receive benefits and are not directly harmed by the changes in the taxes and wages, however, also profitting from an increase in the return-rates. 
Therefore, they do prefer an increase in benefits. 
Corner solutions (i.e. 0 benefit or the full-insurance case) are both not prefered by either group. Note that taxes also affect the future income of the unemployed, since they have a chance to be employed with a certain probability, this thus reduces future income. It works the same way for the employed the other way around.
